---
author: "Wanjiku Githu,"
date: "October 29, 2020"
title: "Analysis of Advertising Data "
output:
  md_document:
    variant: markdown_github
---
**Defining the Question**

The goal is to identify which individuals are most likely to click on ads displayed in a blog based on data that was previously collected on the site.

**Defining the Metrics for Success**

The project will be deemed successful when the characteristics of individuals likely to click ads will be identified through multivariate exploratory data analysis.

**Recording the Experimental Design**

1. Reading in the data
2. Checking for missing values
3. Checking for duplicates
4. Checking for outliers
5. Univariate Exploratory Data Analysis
6. Bivariate Exploratory Data Analysis
7. Multivariate Exploratory Data Analysis
8. Conclusion and Recommendation

**Is the data appropriate to answer the question?**

Yes, the data is suitable for answering the question as the data was collected on the site that will be used to host the cryptography online class advertisement.

**Loading the required libraries**

```{r setup, include=TRUE, warning=FALSE, echo=TRUE, message=FALSE}
# Loading libraries
library(readr)
library(ggplot2)
library(dplyr)
library(tibble)
library(stringr)
library(tidyr)
library(MASS)
library(lubridate)

```

**Reading in the data**
```{r, echo=TRUE, include=TRUE, warning=FALSE, message=FALSE}

# Reading the data
ads <- read_csv(file = 'C:/Users/Githu/Downloads/advertising.csv')
```

**Checking the Data**
```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Determining the number of records
dim(ads)

# Displaying the structure of the dataset
str(ads)

# Viewing the top of our dataset
knitr::kable(
  ads[1:5, ]
  )

# Viewing the bottom of our dataset
knitr::kable(
  ads[995:1000, ] 
  )

# Standardizing the column names
names(ads) <- str_replace_all(names(ads), c(" " = "_"))
names(ads) <- tolower(names(ads))
colnames(ads)

# Checking for missing values
colSums(is.na(ads))

# Checking for duplicates
duplicated_rows <- ads[duplicated(ads),]
duplicated_rows

# Checking for outliers using a boxplot
ads %>% 
  dplyr::select(daily_time_spent_on_site, daily_internet_usage, age, male) %>%
  tidyr::pivot_longer(., cols = c(daily_time_spent_on_site,daily_internet_usage,age),names_to = "Var", values_to = "Val") %>%
  ggplot(aes(x = Var, y = Val, fill = male)) +
  ggtitle("Boxplot of numerical columns") +
  geom_boxplot()

# Boxplot of the area_income variable
boxplot(ads$area_income)
```


The dataset requires little cleaning. There are no missing values or duplicates. The columns names have been standardised to have the same format. Outliers are present in the area income variable only. These are the entries of individuals with low income. We will work with these outliers as dropping them means such individuals with lower incomes(actual potential customers) will not be accounted for.

**Univariate Exploratory Data Analysis**

```{r, echo=TRUE}
# Determining the average age of the visitors
average_age <- mean(ads$age)
paste("The average age of the visitors is", average_age)

# Determining the average income of the visitors
average_income <- mean(ads$area_income)
paste("The average area income of the visitors is:", average_income)

# Determining the average daily time spent on the site
average_time <- mean(ads$daily_time_spent_on_site)
paste("The average daily time spent on the site is:", average_time)

# Determining the average daily internet usage
average_usage <- mean(ads$daily_internet_usage)
paste("The average daily internet usage is:", average_usage)

# Determining the frequency count of male column
# 0 is female and 1 is male
table(ads$male)

# Determining the frequency count of the clicked_ad column
# 0 = did not click ad  and 1 = clicked ad
table(ads$clicked_on_ad)

# Summary statistics of the dataset
summary(ads)

```

**Bivariate Exploratory Data Analysis**

```{r, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
# Bar plot of gender and age
ggplot(data=ads, aes(x=male, y=age)) +
  geom_bar(stat="identity", width=0.5) +
  ggtitle("Barplot of age based on gender")

# Selecting the numerical columns
num_cols <- ads[,c(1, 2, 3, 4)]
colnames(num_cols)

# Correlation matrix of the continous data
cor(num_cols)

# Distribution of area income
hist(ads[ads$clicked_on_ad == 1,]$area_income,xlab = 'Area Income', main = 'Clicked on ads' )
hist(ads[ads$clicked_on_ad == 0,]$area_income,xlab = 'Area Income', main = 'Did not click ads' )

# Average Income of users based on whether they clicked the ads or not
ads %>%
  group_by(clicked_on_ad) %>%
  summarise(mean(area_income))

# Finding the age group that most clicks on ads using a histogram
ads %>%
  filter(clicked_on_ad == 1) %>%
  ggplot(aes(x=age)) +
  geom_histogram(binwidth = 10, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  ggtitle("Age Distribution of the users that clicked the ad") 

```

The age group that most clicked on the ads ranges between 35 and 45

```{r, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
# Finding the age group that did not click on the ads using a histogram
ads %>%
  filter(clicked_on_ad == 0) %>%
  ggplot(aes(x=age)) +
  geom_histogram(binwidth = 10, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  ggtitle("Age Distribution of the users that did not click the ad")
```

Most users who did not click on the ads are between the age of 25 and 35
```{r, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE }
# Distribution of daily time spent on site
ads %>%
  filter(clicked_on_ad == 1) %>%
  ggplot(aes(x=daily_time_spent_on_site)) +
  geom_histogram(binwidth = 10, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  ggtitle("Distribution of daily time spent on site") 

```

Most of the users who clicked on the ads spent 30 to 50 minutes on the site daily

```{r, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
# Histogram of daily internet usage
ads %>%
  filter(clicked_on_ad == 1) %>%
  ggplot(aes(x=daily_internet_usage)) +
  geom_histogram(binwidth = 10, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  ggtitle("Histogram of daily internet usage")
```

Most users who clicked on the ads spent 120 to 150 minutes daily on the internet

```{r, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
# Exploring the timestamp summary
# Plotting the months (clicked_ads = 1)
ads %>%
  filter(clicked_on_ad == 1) %>%
  mutate(ads_months = month(timestamp, label = TRUE)) %>%
  ggplot(aes(x=ads_months)) +
  ggtitle("Months") +
  geom_bar()
```

The month of February recorded the highest number of the clicked ads

```{r, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
# Plotting the weekdays (clicked_ads = 1)
ads %>%
  filter(clicked_on_ad == 1) %>%
  mutate(ads_weekdays = wday(timestamp, label = TRUE)) %>%
  ggplot(aes(x=ads_weekdays)) +
  ggtitle("Weekdays") +
  geom_bar()
```

Most ads were clicked on Sunday, Wednesday or Thursday
```{r, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}

# Exploring the time concept 
# Plotting the hour componenr (clicked_ads = 1)
ads %>%
  filter(clicked_on_ad == 1) %>%
  mutate(ads_hours = hour(timestamp)) %>%
  group_by(ads_hours) %>%
  ggplot(aes(x=ads_hours)) +
  ggtitle("Hour") +
  geom_bar()

```

The highest number of clicked ads were recorded between 9th hour and the lowest at the 10th hour

```{r, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
# Countries 
ads %>%
  dplyr::select(country, clicked_on_ad) %>%
  filter(clicked_on_ad == 1) %>%
  group_by(country) %>%
  summarise(n=n()) %>%
  arrange(desc(n))

ads %>%
  dplyr::select(city, clicked_on_ad, country) %>%
  filter(clicked_on_ad == 1) %>%
  group_by(city) %>%
  summarise(n=n()) %>%
  arrange(desc(n))
```

The countries with relatively higher number of users that clicked the ads include: Australia, Ethiopia, Turkey, Liberia, Liechenstein, South Africa, Afghanistan, France, Hungary and Mayotte.

**Conclusion**

The characteristics of the blog's users that are likely to click on the ads include:

1. The user is female
2. Ages between 35 and 45 years
3. Has an income ranging from 40,000 to 60,000
4. Spends 30 to 50 minutes daily on the site
5. Spends 120 to 150 minutes daily on the internet
6. Visits the blog either on a Sunday, Wednesday or Thursday
7. Visits the blog in early morning hours
8. Their country location is among the following countries:Australia, Ethiopia, Turkey, Liberia, Liechenstein, South Africa, Afghanistan, France, Hungary and Mayotte.


**Recommendations**

Implement the likelihood characteristics mentioned above to enhance targeted advertising.

